A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2018-09-18 doc/large_strain.tex
a bit more on GMG

Efficient and scalable matrix-vector products is not enough to result in a method with linear scaling with respect to the number of unknown DoFs $N$.
It has been widely accepted that the geometric multigrid (GMG) preconditioner can greatly improve the performance of the Conjugate Gradient (CG) solver.
In particular, the number of CG iterations for convergence becomes independent of the number of mesh refinements and thus leads to the $\mathcal O (N)$ solution method.

For this work we adopt the matrix-free level operators in GMG and Chebyshev smoothing. Given a generally heterogeneous non-linear material, definition of transfer operations (restriction and prolongation) and level operators is not trivial. One approach is to start from a arbitrary heterogeneous material at the fine scale and employ non-linear homogenization theory to design specific  transfer operators \cite{Miehe2007}. In this approach the fine-scale displacement is decomposed into long-wave and short-wave contributions by additive split, where the former is associated to the homogeneous response of a patch of elements and the latter represents fluctuations due to the heterogeneous structure of the patch. In case of a homogeneous patch, the transfer operators reduce to the standard geometric transfer. Based on this conclusion we develop our multigrid preconditioner. Namely, we assume that the coarsest level can accurately describe the heterogeneous fine-strain elastic material. In this case the restriction and prolongation operations for patches of elements are always performed for the same material. In order to define the level operators we adopt the following approach. Recall that the fine scale tangent is obtain by linearization around the current displacement field $\gz u^h$. We then restrict this displacement field to multigrid levels and evaluate tangent operators using the matrix-free algorithms from Section \ref{sec:mf}. At the coarsest level we adopt CG solver which shall reduce residual by three orders of magnitude within at most hundred of steps.

@@

We also studies the performance of the GMG preconditioner with standard geometric transfer operations between each level. The GMG was applied to heterogeneous material assuming that the coarsest level can provide adequate discretization of heterogeneity. The numerical studies indicate that the proposed preconditioner lead to the solution approach that is faster than the matrix-based algebraic multigrid preconditioner for moderate polynomial degrees. However the considerable amount of time is spent in the coarse level solver. Therefore our future work will be focused on extending this proposed here matrix-free multigrid solution approach to homogenization based transfer operators \cite{Miehe2007}.

>>>>

Efficient and scalable matrix-vector products is not enough to obtain a method with linear scaling with respect to the number of unknown DoFs $N$.
Efficiency of standard iterative solvers deteriorate for large systems of linear equations.
It has been widely accepted that the geometric multigrid (GMG) preconditioners \cite{Bramble1990, Briggs2000, Janssen2011,May2015} can greatly improve the performance of iterative solvers by smoothing the defect on a hierarchy of levels.
As a result the number of iterations for convergence becomes independent of the number of mesh refinements and thus leads to the $\mathcal O (N)$ solution method. For further information about geometric multigrid methods we refer the reader to \cite{Briggs2000}.

For this work we adopt the matrix-free level operators within the GMG and Chebyshev pre- and post-smoothing on each level. Given a generally heterogeneous non-linear material, definition of transfer operations (restriction and prolongation) and level operators is not trivial. One approach is to start from an arbitrary heterogeneous material at the fine scale and employ non-linear homogenization theory to design specific  transfer operators \cite{Miehe2007}. In this approach the fine-scale displacement is decomposed into long-wave and short-wave contributions by additive split, where the former is associated to the homogeneous response of a patch of elements and the latter represents fluctuations due to the heterogeneous structure of the patch. In the case of a homogeneous patch, the transfer operators reduce to the standard geometric transfer. Based on this observation we adopt the following approach. It is assumed that the mesh at the coarsest level can accurately describe the heterogeneous fine-strain elastic material. In this case the restriction and prolongation operations for patches of elements are always performed for the same material and thus admits usage of standard geometric restriction and prolongation operations. We define the level operators as follows. Recall that the fine scale tangent is obtain by linearization around the current displacement field $\gz u^h$. We then restrict this displacement field to multigrid levels and evaluate tangent operators using the matrix-free algorithms from Section \ref{sec:mf}, essentially corresponding to the linearization around smoothed representation of displacement field.
Note that by doing so the coarse level operator (tangent) $K^{l+1}$ on level $l+1$ is not directly related to the fine level operator (tangent) $K^{l}$ on level $l$ via $K^{l+1}=I^{l+1}_{l}  K^l I^l_{l+1}$, where $I^l_{l+1}$ and $I^{l+1}_l$ are prolongation (coarse-to-fine) and restriction (fine-to-coarse) operators, respectively.
Finally, at the coarsest level we adopt CG solver which shall reduce residual by three orders of magnitude within at most hundred of steps.
The classical multigrid V-cycle approach with two pre- and post-smoothing is adopted.

@@

We also studies the performance of the GMG preconditioner with standard geometric transfer operations between each level. The GMG was applied to heterogeneous material assuming that the coarsest level can provide adequate discretization of heterogeneity. The numerical studies indicate that the proposed preconditioner lead to the solution approach that is faster than the matrix-based algebraic multigrid preconditioner for moderate polynomial degrees. However the considerable amount of time is spent in the coarse level solver. Therefore our future work will be focused on extending this proposed here matrix-free multigrid solution approach to homogenization based transfer operators \cite{Miehe2007}. We are also interested  extending these methods to incompressible three-field formulation.
